---
title: "create_matrices"
author: "Priyesh Verma"
date: "2024-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ==============================
# Load Required Libraries
# ==============================
library(dplyr)
library(readr)
library(lubridate)
library(opentripplanner)
```


```{r}
# ==============================
# SECTION 1: Set Parameters
# ==============================

# Set date and time parameters
year <- 2017
month <- 10
day <- 19
hour <- 8
minute <- 0
second <- 0

# Set maximum travel time (in seconds)
max_time_sec <- 50000

# Set transport modes
modes <- "WALK,BUS,RAIL,TRAM,TRANSIT,SUBWAY"

# Set clamp initial wait time
clamp_initial_wait <- 0
```

Section 1: Set Parameters

- Set the date and time parameters (year, month, day, hour, minute, second).
- Set the maximum travel time (in seconds).
- Set the transport modes.
- Set the clamp initial wait time.

```{r}
# ==============================
# SECTION 2: Load Data
# ==============================

# Load origin and destination points
odpoints <- read_csv("ODpoints_tracts.csv")
```

Section 2: Load Data

- Load the origin and destination points from the ODpoints_tracts.csv file.

```{r}
# ==============================
# SECTION 3: Set Up OpenTripPlanner
# ==============================

# Set up the OTP router
otp <- opentripplanner(
  router = "chicago",
  graphs = ".",
  modes = modes,
  clampInitialWait = clamp_initial_wait
)

# Set the departure time
departure_time <- ymd_hms(paste(year, month, day, hour, minute, second, sep = "-"), tz = "UTC")
```

Section 3: Set Up OpenTripPlanner

- Set up the OTP router with the specified parameters (router name, graphs directory, modes, and clamp initial wait).
- Set the departure time.

```{r}
# ==============================
# SECTION 4: Perform Routing Calculations
# ==============================

# Initialize a data frame to store results
results <- data.frame(
  depart_time = character(),
  origin = character(),
  destination = character(),
  walk_distance = numeric(),
  travel_time = numeric(),
  boardings = numeric()
)

# Iterate over origin points
for (origin in odpoints$GEOID) {
  # Set the origin point
  req <- otp_request(
    otp,
    fromPlace = c(odpoints$X[odpoints$GEOID == origin], odpoints$Y[odpoints$GEOID == origin]),
    dateTime = departure_time,
    maxTimeSec = max_time_sec
  )
  
  # Perform routing calculation
  spt <- otp_plan(req)
  
  # Check if a valid result was found
  if (!is.null(spt)) {
    # Evaluate the SPT for all destination points
    result <- otp_eval(spt, odpoints)
    
    # Add the results to the data frame
    for (r in result) {
      results <- rbind(results, data.frame(
        depart_time = format(departure_time, "%H:%M:%S"),
        origin = r$origin,
        destination = r$destination,
        walk_distance = r$walkDistance,
        travel_time = r$time,
        boardings = r$boardings
      ))
    }
  }
}
```

Section 4: Perform Routing Calculations

- Initialize a data frame to store the results.
- Iterate over origin points.

For each origin point:
-- Set the origin point and create an OTP request.
-- Perform the routing calculation using the otp_plan function.
-- Check if a valid result was found.
-- If a valid result was found, evaluate the SPT for all destination points using the otp_eval function.
-- Add the results to the data frame.

```{r}

# ==============================
# SECTION 5: Save Results
# ==============================

# Save the results as a CSV file
output_file <- paste0("traveltime_matrix_", month, ".", day, "-", hour, ".", minute, ".csv")
write_csv(results, output_file)

# Print the elapsed time
print(paste("Elapsed time was", Sys.time() - start_time, "seconds"))
```

Section 5: Save Results

- Save the results data frame as a CSV file with the filename format traveltime_matrix_<month>.<day>-<hour>.<minute>.csv.
- Print the elapsed time for the entire process.