---
title: "filter_generate"
author: "Priyesh Verma"
date: "2024-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ==============================
# Load Required Libraries
# ==============================
library(dplyr)
library(readr)
```

```{r}
# ==============================
# SECTION 1: Load and Preprocess Data
# ==============================

# Load AVL data for October 19, 2017
oct19 <- read_csv("avl_10_19_17.csv")

# Load GTFS trips data
gtfs_trips <- read_csv("gtfs_trips_oct_17.txt", col_names = TRUE, col_types = cols(
  route_id = col_character(),
  service_id = col_character(),
  trip_id = col_character(),
  schd_trip_id = col_character()
))

# Group AVL data by trip_id
grouped_avl <- group_by(oct19, trip_id)
```

Section 1: Load and Preprocess Data

- Load the AVL data for October 19, 2017, from the avl_10_19_17.csv file.
- Load the GTFS trips data from the gtfs_trips_oct_17.txt file, specifying column types.
- Group the AVL data by trip_id for easier iteration.

```{r}
# ==============================
# SECTION 2: Create trips.csv
# ==============================

# Initialize a data frame to store trip information
avl_trip_df <- data.frame(
  route_id = character(),
  service_id = character(),
  trip_id = character(),
  stringsAsFactors = FALSE
)

# Iterate over each trip_id group in the AVL data
for (trip_id in unique(oct19$trip_id)) {
  group <- filter(grouped_avl, trip_id == !!trip_id)
  
  # Find matching rows in the GTFS trips data based on schd_trip_id
  match_rows <- gtfs_trips %>%
    filter(schd_trip_id == as.character(trip_id))
  
  # If matching rows are found, append the trip information to the data frame
  if (nrow(match_rows) > 0) {
    avl_trip_df <- rbind(avl_trip_df, match_rows[, c("route_id", "service_id", "trip_id")])
  }
}

# Save the trip information as trips.csv
write_csv(avl_trip_df, "trips.csv")
```

Section 2: Create trips.csv

- Initialize a data frame avl_trip_df to store the trip information.
- Iterate over each unique trip_id in the AVL data.
- Find matching rows in the GTFS trips data based on schd_trip_id.
- If matching rows are found, append the trip information to the avl_trip_df data frame.
- Save the avl_trip_df data frame as trips.csv.

```{r}
# ==============================
# SECTION 3: Create stop_times.csv
# ==============================

# Load valid stop IDs from GTFS stops.txt
valid_stops <- read_csv('gtfs-10-22-17/stops.txt')
gtfs_stops <- unique(valid_stops$stop_id)

# Initialize a data frame to store stop time information
stop_times <- data.frame(
  trip_id = character(),
  arrival_time = character(),
  departure_time = character(),
  stop_id = numeric(),
  stop_sequence = numeric(),
  stringsAsFactors = FALSE
)

# Iterate over each trip_id group in the AVL data
for (trip_id in unique(oct19$trip_id)) {
  group <- filter(grouped_avl, trip_id == !!trip_id)
  
  # Find matching rows in the GTFS trips data based on schd_trip_id
  match_rows <- gtfs_trips %>%
    filter(schd_trip_id == as.character(trip_id))
  
  # If matching rows are found
  if (nrow(match_rows) > 0) {
    # Iterate over each row in the AVL data for the current trip_id
    for (i in seq_len(nrow(group))) {
      time_update <- group[i, ]
      
      # Check if the stop_id is a valid GTFS stop_id
      if (time_update$stop_id %in% gtfs_stops) {
        # Append the stop time information to the data frame
        stop_times <- rbind(stop_times, data.frame(
          trip_id = match_rows$trip_id[1],
          arrival_time = time_update$time,
          departure_time = time_update$time,
          stop_id = time_update$stop_id,
          stop_sequence = time_update$stop_sequence
        ))
      }
    }
  }
}

# Save the stop time information as stop_times.csv
write_csv(stop_times, "stop_times.csv")
```

Section 3: Create stop_times.csv

- Load valid stop IDs from the GTFS stops.txt file.
- Initialize a data frame stop_times to store the stop time information.
- Iterate over each unique trip_id in the AVL data.
- Find matching rows in the GTFS trips data based on schd_trip_id.
- If matching rows are found, iterate over each row in the AVL data for the current trip_id.
- Check if the stop_id is a valid GTFS stop_id.
- If it is a valid stop_id, append the stop time information to the stop_times data frame.
- Save the stop_times data frame as stop_times.csv.

```{r}
# ==============================
# SECTION 4: Include Rail Trips and Stop Times
# ==============================

# Filter rail routes from GTFS trips data
rail <- gtfs_trips %>%
  filter(route_id %in% c("Red", "Blue", "Brn", "G", "Org", "Pink", "P", "Y")) %>%
  select(-schd_trip_id)

# Append rail trips to the avl_trip_df data frame
avl_trip_df <- rbind(avl_trip_df, rail)

# Save the updated trip information as trips.csv
write_csv(avl_trip_df, "trips.csv")

# Load original GTFS stop_times data for October 19
original_stopT <- read_csv('gtfs_oct19_stop_times.txt', col_names = TRUE, col_types = cols(
  trip_id = col_character(),
  arrival_time = col_character(),
  departure_time = col_character(),
  stop_id = col_double(),
  stop_sequence = col_double()
))

# Iterate over each row in the rail trips data
for (i in seq_len(nrow(rail))) {
  r <- rail[i, ]
  
  # Find matching rows in the original GTFS stop_times data based on trip_id
  new <- original_stopT %>%
    filter(trip_id == r$trip_id)
  
  # Append the matching stop_times rows to the stop_times data frame
  stop_times <- rbind(stop_times, new)
}

# Save the updated stop_times data to stop_times.csv
write_csv(stop_times, "stop_times.csv")
```

Section 4: Include Rail Trips and Stop Times

- Filter rail routes from the GTFS trips data based on the route_id.
- Append the rail trips to the avl_trip_df data frame.
- Save the updated avl_trip_df data frame as trips.csv.
- Load the original GTFS stop_times data for October 19 from the gtfs_oct19_stop_times.txt file, specifying column types.
- Iterate over each row in the rail trips data.
- Find matching rows in the original GTFS stop_times data based on trip_id.
- Append the matching stop_times rows to the stop_times data frame.
- Save the updated stop_times data frame as stop_times.csv.