---
title: "get_travel_time"
author: "Priyesh Verma"
date: "2024-03-17"
output: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Importing Required Libraries
library(dplyr)
library(sf)
library(lubridate)
library(sqldf)
```


```{r}
# Section 1: Data Import and Preprocessing
# Import data
input_origins <- read.csv("input_origins.csv")
input_dests <- read.csv("input_dests.csv")
odx_data <- read.csv("1019_trips_filtered.csv")
stops_geo <- read.csv("./ODX_files/PT_2017-10-10_stops.csv")
census_tracts <- read.csv("census_tracts_bounded.csv")

# Convert character columns to appropriate data types
odx_data$dep_time <- as.POSIXct(odx_data$dep_time, format = "%Y-%m-%d %H:%M:%S")
odx_data$arr_time <- as.POSIXct(odx_data$arr_time, format = "%Y-%m-%d %H:%M:%S")
```

Section 1: Data Import and Preprocessing

This section imports the required data files and performs necessary data type conversions.

```{r}
# Section 2: Create and Connect to Local SQLite Database of ODX Trip Data
# Create a SQLite database connection
con <- dbConnect(SQLite(), dbname = "od_trips.db")

# Create a table to store ODX trip data
dbExecute(con, "CREATE TABLE IF NOT EXISTS od_trips (
  origin INTEGER,
  dest INTEGER,
  dep_time TIMESTAMP,
  arr_time TIMESTAMP,
  transfers INTEGER,
  travel_time INTEGER,
  UNIQUE (origin, dest, dep_time, arr_time, transfers, travel_time)
)")

# Insert ODX trip data into the database
odx_data_rows <- apply(odx_data, 1, function(row) {
  dbExecute(con, "INSERT OR IGNORE INTO od_trips VALUES (?, ?, ?, ?, ?, ?)",
            params = as.list(row))
})

# Commit changes and close the database connection
dbCommit(con)
dbDisconnect(con)
```

Section 2: Create and Connect to Local SQLite Database of ODX Trip Data

- This section creates a SQLite database and a table to store the ODX trip data.
- The ODX trip data is inserted into the database table.

```{r}
# Section 3: Add Census Tract Geoid to Input OD Pairs
# Convert census tract geometries to SF objects
census_tracts_sf <- st_as_sf(census_tracts, wkt = "geometry")

# Function to find the census tract for a given coordinate
find_census_tract <- function(lon, lat) {
  pt <- st_point(c(lon, lat))
  tract_filter <- st_intersects(census_tracts_sf, pt, sparse = FALSE)
  if (any(tract_filter)) {
    census_tracts_sf$geoid10[tract_filter][1]
  } else {
    NA_integer_
  }
}

# Add census tract Geoid to input OD pairs
input_origins$start_tract <- vapply(
  seq_len(nrow(input_origins)),
  function(i) find_census_tract(input_origins$X[i], input_origins$Y[i]),
  FUN.VALUE = integer(1)
)

input_dests$end_tract <- vapply(
  seq_len(nrow(input_dests)),
  function(i) find_census_tract(input_dests$X[i], input_dests$Y[i]),
  FUN.VALUE = integer(1)
)
```

Section 3: Add Census Tract Geoid to Input OD Pairs

- This section converts the census tract geometries to SF objects.
- A function find_census_tract is defined to find the census tract for a given coordinate.
- The census tract Geoid is added to the input OD pairs dataframes.

```{r}
# Section 4: Supporting Function to Calculate Lon/Lat Bounds for a Census Tract
bounds <- function(tract_geoid) {
  tract <- census_tracts[census_tracts$geoid10 == tract_geoid, ]
  if (nrow(tract) == 0) {
    return(NULL)
  }
  c(tract$minx, tract$maxx, tract$miny, tract$maxy)
}
```

Section 4: Supporting Function to Calculate Lon/Lat Bounds for a Census Tract

This section defines a function bounds to calculate the longitude and latitude bounds for a given census tract Geoid.

```{r}
# Section 5: Create and Connect to Local SQLite Database of Transit Stops
# Create a SQLite database connection
con_s <- dbConnect(SQLite(), dbname = "stops.db")

# Create a table to store transit stop data
dbExecute(con_s, "CREATE TABLE IF NOT EXISTS stops (
  stop_id INTEGER UNIQUE,
  stop_lat REAL,
  stop_lon REAL
)")

# Insert transit stop data into the database
stops_geo_rows <- apply(stops_geo, 1, function(row) {
  dbExecute(con_s, "INSERT OR IGNORE INTO stops VALUES (?, ?, ?)",
            params = as.list(row[c("stop_id", "stop_lat", "stop_lon")]))
})

# Commit changes and close the database connection
dbCommit(con_s)
dbDisconnect(con_s)
```

Section 5: Create and Connect to Local SQLite Database of Transit Stops

- This section creates a SQLite database and a table to store the transit stop data.
- The transit stop data is inserted into the database table.

```{r}
# Section 6: OD Pair to ODX Trip Matching
# Create a SQLite database connection for results
con_results <- dbConnect(SQLite(), dbname = "results.db")

# Create a table to store travel time results
dbExecute(con_results, "CREATE TABLE IF NOT EXISTS results_travel_time (
  time_start INTEGER,
  time_end INTEGER,
  start_lat REAL,
  start_lon REAL,
  end_lat REAL,
  end_lon REAL,
  start_stop INTEGER,
  end_stop INTEGER,
  transfers INTEGER,
  median_tt INTEGER,
  quartile_tt INTEGER,
  UNIQUE (time_start, time_end, start_lat, start_lon, end_lat, end_lon, start_stop, end_stop, transfers, median_tt, quartile_tt)
)")

# Iterate through each desired Origin-Destination pair
for (i in seq_len(nrow(input_origins))) {
  print(i)

  # Extract origin and destination lat/lon in geo coordinates
  or_lat <- input_origins$Y[i]
  or_lon <- input_origins$X[i]
  dest_lat <- input_dests$Y[i]
  dest_lon <- input_dests$X[i]
  start_id <- input_origins$start_tract[i]
  end_id <- input_dests$end_tract[i]

  # Get min&max bounds on lat/lon given desired radius
  origin_bounds <- bounds(start_id)
  dest_bounds <- bounds(end_id)

  # Case where origin or destination coordinates are not in a census tract
  if (is.null(origin_bounds) || is.null(dest_bounds)) {
    dbExecute(con_results, "INSERT OR IGNORE INTO results_travel_time VALUES (?, ?, ?, ?, ?, ?, ?, ?, ?, ?, ?)",
              params = list(28800, 28800, or_lat, or_lon, dest_lat, dest_lon, NA, NA, NA, NA, NA))
    next
  }

  # Get a list of stops within census tracts
  or_neighbors <- dbGetQuery(con_s, "SELECT * FROM stops WHERE
                              stop_lon > ? AND stop_lon < ? AND
                              stop_lat > ? AND stop_lat < ?",
                              params = origin_bounds)

  dest_neighbors <- dbGetQuery(con_s, "SELECT * FROM stops WHERE
                                stop_lon > ? AND stop_lon < ? AND
                                stop_lat > ? AND stop_lat < ?",
                                params = dest_bounds)

  # Find all possible trips in ODX database that match neighboring origin and destination id
  matching_trips <- dbGetQuery(con, "SELECT * FROM od_trips WHERE
                                origin IN (?) AND dest IN (?)",
                                params = list(or_neighbors$stop_id, dest_neighbors$stop_id))

  if (nrow(matching_trips) > 0) {
    # Filter trips for morning rush hour (8 AM +/- 2 hours)
    matching_trips <- matching_trips[abs(hour(matching_trips$dep_time) - 8) < 2, ]

    # Sort matching trips by travel time
    matching_trips <- matching_trips[order(matching_trips$travel_time), ]

    # Calculate 75th percentile and median travel time indices
    quartile_index <- ceiling(0.75 * nrow(matching_trips))
    median_index <- ceiling(0.5 * nrow(matching_trips))

    # Get the 75th percentile and median travel time trips
    q_trip <- matching_trips[quartile_index, ]
    m_trip <- matching_trips[median_index, ]

    # Insert results into the results_travel_time table
    dbExecute(con_results, "INSERT OR IGNORE INTO results_travel_time VALUES (?, ?, ?, ?, ?, ?, ?, ?, ?, ?, ?)",
              params = list(q_trip$dep_time, q_trip$arr_time, or_lat, or_lon, dest_lat, dest_lon,
                            q_trip$origin, q_trip$dest, q_trip$transfers, m_trip$travel_time, q_trip$travel_time))
  }

  dbCommit(con_results)
}

# Visualize results
res <- dbGetQuery(con_results, "SELECT * FROM results_travel_time")
print(res)

# Close database connections
dbDisconnect(con_results)
dbDisconnect(con)
dbDisconnect(con_s)
```

Section 6: OD Pair to ODX Trip Matching
- This section creates a SQLite database and a table to store the travel time results.

It iterates through each desired Origin-Destination pair and performs the following steps:
- Extract origin and destination lat/lon coordinates.
- Get the min&max bounds on lat/lon for the origin and destination census tracts.
- Get a list of stops within the origin and destination census tracts.
- Find all possible trips in the ODX database that match the neighboring origin and destination stops.
- Filter trips for morning rush hour (8 AM +/- 2 hours).
- Sort matching trips by travel time and calculate the 75th percentile and median travel time.
- Insert the results into the results_travel_time table.
