---
title: "random_subsets"
author: "Priyesh Verma"
date: "2024-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Section 1: Load Required Libraries
library(readr)
library(dplyr)
```

Section 1: Load Required Libraries

This section loads the required libraries: readr for reading CSV files and dplyr for data manipulation.

```{r}
# Section 2: Read Census Tracts Origin Data
census_tracts <- "ODpoints_tracts.csv"
ODtracts_df <- read_csv(census_tracts)
```

Section 2: Read Census Tracts Origin Data

This section reads the census tracts origin data from the "ODpoints_tracts.csv" file using the read_csv function from the readr package and stores it in the ODtracts_df data frame.

```{r}
# Section 3: Sample and Write to CSV
for (percentage in seq(10, 100, 10)) {
 sample_percentage <- percentage / 100
 mysample <- ODtracts_df %>%
   sample_frac(sample_percentage)
 
 output_file <- sprintf("%d%%ODpoints.csv", percentage)
 write_csv(mysample, output_file, col_names = TRUE)
}
```

Section 3: Sample and Write to CSV

This section uses a for loop to iterate over percentages from 10% to 100% in increments of 10%.

Inside the loop:
- The sample_percentage is calculated by dividing the current percentage by 100.
- The sample_frac function from dplyr is used to sample a fraction of the ODtracts_df data frame based on the sample_percentage.
- The sprintf function is used to generate the output file name in the format "%d%%ODpoints.csv", where %d is replaced with the current percentage value.
- The sampled data frame is written to the output file using the write_csv function from the readr package, with column names included (col_names = TRUE).
