---
title: "travel_time"
author: "Priyesh Verma"
date: "2024-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ==============================
# Load Required Libraries
# ==============================
library(dplyr)
library(readr)
library(lubridate)
```

```{r}
# ==============================
# SECTION 1: Set Parameters
# ==============================

# Set date and time parameters
year <- 2017
month <- 10
day <- 19
hour <- 8
minute <- 0
second <- 0

# Set maximum travel time (in seconds)
max_time_sec <- 50000

# Set transport modes
modes <- "WALK,BUS,RAIL,TRAM,TRANSIT,SUBWAY"

# Set clamp initial wait time
clamp_initial_wait <- 0
```

Section 1: Set Parameters

- Set the date and time parameters (year, month, day, hour, minute, second).
- Set the maximum travel time (in seconds).
- Set the transport modes.
- Set the clamp initial wait time.

```{r}
# ==============================
# SECTION 2: Load Data
# ==============================

# Load origin points
input_origins <- read_csv("input_origins.csv")

# Load destination points
input_dests <- read_csv("input_dests.csv")
```

Section 2: Load Data

- Load the origin points from the input_origins.csv file.
- Load the destination points from the input_dests.csv file.

```{r}
# ==============================
# SECTION 3: Process Origin-Destination Pairs
# ==============================

# Create a list of origin and destination points
origin_list <- input_origins
dest_list <- input_dests

# Initialize a data frame to store results
results <- data.frame(
  date = character(),
  depart_time = character(),
  origin = character(),
  destination = character(),
  avl_found = logical(),
  walk_distance = numeric(),
  travel_time = numeric(),
  boardings = numeric()
)

# Set the departure time
departure_time <- ymd_hms(paste(year, month, day, hour, minute, second, sep = "-"))

# Iterate over origin-destination pairs
for (i in seq_along(origin_list)) {
  print(paste("Processing origin:", i))
  
  # Set the origin point
  origin <- origin_list[i, ]
  
  # Set the destination point
  destination <- dest_list[i, ]
```

```{r}
# ==============================
# Perform Routing Calculation
# ==============================

# Load the OTP package
library(opentripplanner)

# Set up the OTP router
otp <- opentripplanner(
  router = "chicago",
  graphs = ".",
  modes = modes,
  clampInitialWait = clamp_initial_wait
)

# Set the departure time
departure_time <- as.POSIXct(departure_time, tz = "UTC")

# Iterate over origin-destination pairs
for (i in seq_along(origin_list)) {
  print(paste("Processing origin:", i))
  
  # Set the origin point
  origin <- origin_list[i, ]
  
  # Set the destination point
  destination <- dest_list[i, ]
  
  # Perform routing calculation
  result <- otp_plan(
    otp,
    fromPlace = c(origin$X, origin$Y),
    toPlace = c(destination$X, destination$Y),
    dateTime = departure_time,
    maxTimeSec = max_time_sec
  )
  
  # Check if a valid result was found
  avl_found <- !is.null(result)
  
  # Extract relevant information from the result
  if (avl_found) {
    walk_distance <- result$walkDistance
    travel_time <- result$time
    boardings <- result$boardings
  } else {
    walk_distance <- NA
    travel_time <- NA
    boardings <- NA
  }
  
  # Add the result to the data frame
  results <- rbind(results, data.frame(
    date = format(departure_time, "%m/%d/%Y"),
    depart_time = format(departure_time, "%H:%M:%S"),
    origin = origin$GEOID,
    destination = destination$GEOID,
    avl_found = avl_found,
    walk_distance = walk_distance,
    travel_time = travel_time,
    boardings = boardings
  ))
}
```

- Load the opentripplanner package, which provides an R interface to the OpenTripPlanner routing engine.
- Set up the OTP router with the specified parameters (router name, graphs directory, modes, and clamp initial wait).
- Convert the departure time to a POSIXct object for use with the OTP functions.
- Iterate over the origin-destination pairs.

For each origin-destination pair:
-- Set the origin and destination points from the input data.
-- Perform the routing calculation using the otp_plan function, passing the origin and destination coordinates, departure time, and maximum travel time.
-- Check if a valid result was found.
-- If a valid result was found, extract the walk distance, travel time, and number of boardings from the result.
-- Add the result to the results data frame.

```{r}
# Assuming the routing calculation provides the following values:
  avl_found <- TRUE
  walk_distance <- 500
  travel_time <- 3600
  boardings <- 2
  
  # Add the result to the data frame
  results <- rbind(results, data.frame(
    date = format(departure_time, "%m/%d/%Y"),
    depart_time = format(departure_time, "%H:%M:%S"),
    origin = origin$GEOID,
    destination = destination$GEOID,
    avl_found = avl_found,
    walk_distance = walk_distance,
    travel_time = travel_time,
    boardings = boardings
  ))
}
```

Section 3: Process Origin-Destination Pairs

- Create a list of origin and destination points.
- Initialize a data frame to store the results.
- Set the departure time.
- Iterate over origin-destination pairs.
- Set the origin and destination points for each iteration.
- Perform the routing calculation (code for routing calculation goes here).
- Assuming the routing calculation provides the following values: avl_found, walk_distance, travel_time, and boardings.
- Add the result to the data frame.

```{r}
# ==============================
# SECTION 4: Save Results
# ==============================

# Save the results as a CSV file
output_file <- paste0("jobs_avl_traveltime_", month, ".", day, "-", hour, ".", minute, ".csv")
write_csv(results, output_file)
```

Section 4: Save Results

- Save the results data frame as a CSV file with the filename format jobs_avl_traveltime_<month>.<day>-<hour>.<minute>.csv.
